---
title: "Analyses: Disentangling social-relational classroom climate and its relation to students' behavioral problems"
author: "Jürgen Wilbert"
date: "Version (`r Sys.Date()`)"
output:
  html_document:
    theme: cosmo
    df_print: paged
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: false
bibliography: packages.bib
geometry: a4paper
link-citations: yes
biblio-style: apalike
---

```{r setup, include=FALSE}
# load packages -----------------------------------------------------------

if (!("wmisc" %in% installed.packages())) 
  devtools::install_github("jazznbass/wmisc")
if (!("scaledic" %in% installed.packages())) 
  devtools::install_github("jazznbass/scaledic")

packages <- c(
  'sjPlot', 'knitr', 'semTable', 'kableExtra', 'bookdown', 
  'scaledic', 'Amelia', 'wmisc', 'psych', 'lavaan', 'nlme', 'lme4',
  'tidyverse', 'readxl', 'gridExtra', 'broom', 'sjPlot', 'mclust'
)

invisible(sapply(packages, library, character.only = TRUE))

# load data

dat <- readRDS("data_l1.rds")
dat_agg <- readRDS("data_l2.rds")

# Define variables: classclimate/wellbeing/behavioral problems/control for all analyzes
var_cc_l1 <- c("feess_ga", "feess_si", "soc_relatedness","soc_positivity", "sar_victim", "sar_perpetrator")
var_cc_l2 <- paste0(var_cc_l1, "_mean")
var_cc_l2_sd <- paste0(var_cc_l1, "_sd")

var_bp_l1 <- c("itrf_sw", "itrf_ad", "itrf_apd", "itrf_opp")
var_bp_l2 <- paste0(var_bp_l1, "_mean")
var_bp_l2_sd <- paste0(var_bp_l1, "_sd")

var_cont_l1 <- c("age", "grade", "sex", "migration_background")
var_cont_l2 <- paste0(var_cont_l1, "_mean")
var_cont_l2_sd <- paste0(var_cont_l1, "_sd")

# set parameters

knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# function for nicer output
html_table <- function(table, title = "", footnote = "") {
  title <- paste0("Table.<br><i>", title, "</i>")
  kable(table, caption = title) %>%
   kable_classic(full_width = FALSE) %>%
   footnote(footnote)
}

html_regression <- function(fit, ...) {
  sjPlot::tab_model(fit, show.std = TRUE, show.se = TRUE, show.ci = FALSE, ...)
}

```

# Software acknowledgment

The following packages were used for this report:

`R` at version `r R.version$version.string` [@R-base].

```{r package_reference, results="asis", echo=FALSE}
wmisc::reference_package_version(packages)
```

------------------------------------------------------------------------

# Research questions

Q1: Is there a general factor that could be derived from various aspects of students relations within the classroom (ie perceived social inclusion; social position; bullying involvement; feeling of acceptance by the teacher) ? 

Q1a: On an individual level: Is it possible to conflate various aspects of students relations within the classroom into one distinct general factor (social-relational situation)?

Q1b: On a classroom level: Is it possible to conflate the classroom means of various aspects of students relations into on distinct general factor (social-relational classroom climate)?

Q2: What is the connection between students behavioral problems and various aspects of their social relations?

Q2a: On an individual level: To what extend can students behavioral problems in a class be explained by means of various aspects of their social relations?

Q2b: On a classroom level: To what extend can the average amount of behavioral problems in a classroom be explained by classroom means of various aspects of students social relations? 

Q2c: To what extend do classroom means of various aspects of students social relations contribute to individual behavioral classroom problems?

Add on questions(probably they will be dropped later :-) )

> Q3a: Can we find profiles of components of social-emotional wellbeing?
> Q3b: Can we find profiles of components of social-emotional classclimate?
> Q3c: What is the nexus between the profiles of social-emotional wellbeing and behavioral problems?
> Q3c: What is the nexus between the profiles of social-emotional classclimate and behavioral problems?


# Descriptives of all variables

```{r descriptives_l1}
.var <- c(var_cc_l1, var_bp_l1, "age", "sex_male", "migration_background_numeric")

dat %>% select(!!.var) %>% rename_items() %>% descriptives() %>% html_table

dat %>% group_by(grade) %>% summarise(n = n()) %>% html_table
dat %>% group_by(sex) %>% summarise(n = n()) %>% html_table
dat %>% group_by(migration_background) %>% summarise(n = n()) %>% html_table

dat %>% select(!!.var) %>% rename_items() %>% wmisc::nice_corrmatrix(upper = FALSE, numbered_columns = TRUE, char_p10 = "T") %>% html_table
```

```{r descriptives_l2}
.var <- c(var_cc_l2, var_bp_l2, "age_mean", "sex_male_mean", "migration_background_numeric_mean")
dat_agg %>% 
  select(!!.var) %>% 
  rename_items() %>% 
  descriptives() %>% 
  html_table()

dat_agg %>% 
  select(!!.var) %>% rename_items() %>% 
  wmisc::nice_corrmatrix(upper = FALSE, numbered_columns = TRUE, char_p10 = "T") %>% html_table
```


# Q1a

We conducted a parallel analysis to determine the number of factors to be extracted from the data. The analysis point to two factors (see figure xxx)

```{r parallel_l1, fig.cap="Parallel analysis scree plot."}
dat %>% select(!!var_cc_l1) %>% fa.parallel(main = "", fa = "fa")
```

```{r efa_l1}
dat %>% 
  select(!!var_cc_l1) %>% 
  rename_items() %>% 
  exploratory_fa(nfactors = 2, cut = 0.2, fm = "minres") %>% 
  html_table("Exploratory factor analyzes of wellbeing variables.")
```

# Q1b

We conducted a parallel analysis to determine the number of factors to be extracted from the data. The analysis point to two factors (see figure xxx).

```{r nfactors_l2, , fig.cap="Parallel analysis scree plot."}
dat_agg %>% 
  select(!!var_cc_l2) %>% 
  fa.parallel(main = "", fa = "fa")
```

```{r efa_l2}
dat_agg %>% 
  select(!!var_cc_l2) %>% 
  rename_items() %>% 
  exploratory_fa(nfactors = 2, fm = "minres") %>%
  html_table("Exploratory factor analyzes of classclimate variables.")
```

# Q2a

```{r}
fit <- lme(itrf_sw ~ feess_si + feess_ga + soc_positivity + soc_relatedness + sar_victim + sar_perpetrator + age + migration_background + sex, random =~ 1 | id_class_teacher, data = dat, na.action = na.omit)
html_regression(fit, title = "Social withdrawal regressed on wellbing variables")
```

```{r}
fit <- lme(itrf_ad ~ feess_si + feess_ga + soc_positivity  + soc_relatedness + sar_victim + sar_perpetrator+ age + migration_background + sex, random =~ 1 | id_class_teacher, data = dat, na.action = na.omit)
html_regression(fit, title = "Anxiety/depression regressed on wellbing variables")
```

```{r}
fit <- lme(itrf_apd ~ feess_si + feess_ga + soc_positivity  + soc_relatedness + sar_victim + sar_perpetrator+ age + migration_background + sex, random =~ 1 | id_class_teacher, data = dat, na.action = na.omit)
html_regression(fit, title = "Academic disorganized behavior regressed on wellbing variables")
```

```{r}
fit <- lme(itrf_opp ~ feess_si + feess_ga +  soc_positivity  + soc_relatedness + sar_victim + sar_perpetrator+ age + migration_background + sex, random =~ 1 | id_class_teacher, data = dat, na.action = na.omit)
html_regression(fit, title = "Oppositional behavior regressed on wellbing variables")
```

# Q2b

```{r}
fit <- lm(itrf_sw_mean ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean   + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean, data = dat_agg, na.action = na.omit)
html_regression(fit, title = "Social withdrawal regressed on classclimate")
```

```{r}
fit <- lm(itrf_ad_mean ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean + sar_victim_mean + sar_perpetrator_mean+ age_mean + migration_background_numeric_mean + sex_male_mean, data = dat_agg, na.action = na.omit)
html_regression(fit, title = "Anxiety/depression regressed on classclimate")
```

```{r}
fit <- lm(itrf_apd_mean ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean + sar_victim_mean + sar_perpetrator_mean+ age_mean + migration_background_numeric_mean + sex_male_mean, data = dat_agg, na.action = na.omit)
html_regression(fit, title = "Academic disorganized behavior regressed on classclimate")
```

```{r}
fit <- lm(itrf_opp_mean ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean, data = dat_agg, na.action = na.omit)
html_regression(fit, title = "Opositional behavior regressed on wellbing variables")
```

# Q2c

```{r}
dat_l12 <- full_join(dat, dat_agg, by = "id_class_teacher")
```


```{r}
fit <- lme(itrf_sw ~ feess_si + feess_ga + soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age + migration_background + age_mean + sex + migration_background_numeric_mean + sex_male_mean, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)
html_regression(fit, title = "Social withdrawal regressed on wellbing and classclimate variables")
```

```{r}
fit <- lme(itrf_ad ~ feess_si + feess_ga +  soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age + migration_background + age_mean + sex + migration_background_numeric_mean + sex_male_mean, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)
html_regression(fit, title = "Anxiety/Depression regressed on wellbing and classclimate variables")
```

```{r}
fit <- lme(itrf_apd ~ feess_si + feess_ga + soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age + migration_background + age_mean + sex + migration_background_numeric_mean + sex_male_mean, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)
html_regression(fit, title = "Academic disengagement regressed on wellbing and classclimate variables")
```

```{r}
fit <- lme(itrf_opp ~ feess_si + feess_ga + soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age + migration_background + age_mean + sex + migration_background_numeric_mean + sex_male_mean, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)
html_regression(fit, title = "Oppositional behavior regressed on wellbing and classclimate variables")
```


# Q3a

```{r}
dat2 <- dat %>%
  select(!!var_cc_l1) %>%
  rename_items() %>%
  na.omit() %>%
  mutate(across(everything(), scale))

# fit_bic <- mclustBIC(dat2)
# saveRDS(fit_bic, "fit_bic.rds")
fit_bic <- readRDS("fit_bic.rds")

#plot(fit_bic)
#summary(fit_bic)

# fit_icl <- mclustICL(dat2)
# saveRDS(fit_icl, "fit_icl.rds")
# fit_icl <- readRDS("fit_icl.rds")

#plot(fit_icl)
#summary(fit_icl)

# bootstrap <- mclustBootstrapLRT(dat2, modelName = "VEV", maxG = 9)
# saveRDS(bootstrap, "bootstrap.rds")
# bootstrap <- readRDS("bootstrap.rds")
```


```{r}
mod1 <- Mclust(dat2, modelNames = "VEV", G = 3, x = fit_bic)
#summary(mod1)

means <- as.data.frame(mod1$parameters$mean) %>%
  rownames_to_column("Variable") %>%
  pivot_longer(cols = -1) %>%
  mutate(name = recode(name,
    V1 = "Bully and bullied; socially rejected 17%",
    V2 = "Socially included 60%",
    V3 = "Bullying victim; socially rejected 23%"
  ))

p <- means %>%
  ggplot(aes(Variable, value, group = name, color = name)) +
  geom_point(size = 2.25) +
  geom_line(size = 1.25) +
  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p


mod1 <- Mclust(dat2, modelNames = "VEV", G = 5, x = fit_bic)
#summary(mod1)

means <- as.data.frame(mod1$parameters$mean) %>%
  rownames_to_column("Variable") %>%
  pivot_longer(cols = -1) %>%
   mutate(name = recode(name,
    V1 = "Bullying victim; socially rejected 19%",
    V2 = "Socially included; few relations 31%",
    V3 = "Bully and bullied; socially rejected 14%",
    V4 = "Slight bully; slight socially rejected 14%",
    V5 = "Socially included; many relations 22%"
    
  ))

  

p <- means %>%
  ggplot(aes(Variable, value, group = name, color = name)) +
  geom_point(size = 2.25) +
  geom_line(size = 1.25) +
  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p

```

# Q3b

```{r}
dat2 <- dat_agg %>%
  select(!!var_cc_l2) %>%
  rename_items() %>%
  na.omit() %>%
  mutate(across(everything(), scale))

# fit_bic <- mclustBIC(dat2)
# saveRDS(fit_bic, "fit_bic_l2.rds")
fit_bic <- readRDS("fit_bic_l2.rds")

#plot(fit_bic)
#summary(fit_bic)

#bootstrap <- mclustBootstrapLRT(dat2, modelName = "EVE", maxG = 4)
#saveRDS(bootstrap, "bootstrap_l2.rds")
#bootstrap <- readRDS("bootstrap_l2.rds")
```



```{r}
mod1 <- Mclust(dat2, modelNames = "EVE", G = 3, x = fit_bic)
#summary(mod1)

means <- as.data.frame(mod1$parameters$mean) %>%
  rownames_to_column("Variable") %>%
  pivot_longer(cols = -1) %>%
  mutate(name = recode(name,
    V1 = "Average inclusion, lower bullying 70%",
    V2 = "Intense bullying 20%",
    V3 = "High social inclusion 10%"
  ))

p <- means %>%
  ggplot(aes(Variable, value, group = name, color = name)) +
  geom_point(size = 2.25) +
  geom_line(size = 1.25) +
  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p

```

# Q3c

```{r}
dat2 <- dat %>%
  select(!!var_cc_l1) %>% 
  rename_items() %>%
  na.omit() %>%
  mutate(across(everything(), scale))

fit_bic <- readRDS("fit_bic.rds")

mod1 <- Mclust(dat2, modelNames = "VEV", G = 5, x = fit_bic)

dat2 <- dat
dat2$filter <- apply(dat2[,var_cc_l1], 1, function(x) any(is.na(x)))
dat2 <- dat2 %>% filter(!filter)
dat2$profile <- factor(mod1$classification, labels = c(
  "Bullying victim; socially rejected 19%",
  "Socially included; few relations 31%",
  "Bully and bullied; socially rejected 14%",
  "Slight bully; slight socially rejected 14%",
  "Socially included; many relations 22%"))

dat2 <- dat2 %>% mutate(across(!!c(var_bp_l1, var_cc_l1), scale))

dat2 %>% group_by(profile) %>% 
  summarise(across(!!var_bp_l1, mean, na.rm = TRUE)) %>%
  pivot_longer(col = -1) %>%
  ggplot(aes(x = profile, y = value, fill = name)) +
  geom_bar(stat = "identity",position = "dodge" ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


   

```

# Q3d

```{r}
dat2 <- dat_agg %>%
  select(!!var_cc_l2) %>%
  rename_items() %>%
  na.omit() %>%
  mutate(across(everything(), scale))

fit_bic <- readRDS("fit_bic_l2.rds")

mod1 <- Mclust(dat2, modelNames = "EVE", G = 3, x = fit_bic)

dat2 <- dat_agg
dat2$filter <- apply(dat2[,var_cc_l2], 1, function(x) any(is.na(x)))

dat2 <- dat2 %>% filter(!filter)
dat2$profile <- factor(mod1$classification, labels = c("Average inclusion, lower bullying 70%","Intense bullying 20%", "High social inclusion 10%"))


dat2 <- dat2 %>% mutate(across(!!c(var_bp_l2, var_cc_l2), scale))

dat2 %>% group_by(profile) %>% 
  summarise(across(!!var_bp_l2, mean, na.rm = TRUE)) %>%
  pivot_longer(col = -1) %>%
  ggplot(aes(x = profile, y = value, fill = name)) +
  geom_bar(stat = "identity",position = "dodge" )  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Reference
