---
title: "Analyses: Disentangling social-relational classroom climate and its relation to students' behavioral problems"
author: "Jürgen Wilbert"
date: "Version (`r Sys.Date()`)"
output:
  html_document:
    theme: cosmo
    df_print: paged
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: false
bibliography: packages.bib
geometry: a4paper
link-citations: yes
biblio-style: apalike
---

```{r setup, include=FALSE}
# load packages -----------------------------------------------------------

if (!("wmisc" %in% installed.packages())) 
  devtools::install_github("jazznbass/wmisc")
if (!("scaledic" %in% installed.packages())) 
  devtools::install_github("jazznbass/scaledic")

packages <- c(
  'sjPlot', 'knitr', 'semTable', 'kableExtra', 'bookdown', 
  'scaledic', 'Amelia', 'wmisc', 'psych', 'lavaan', 'nlme', 'lme4',
  'tidyverse', 'readxl', 'gridExtra', 'broom', 'sjPlot', 'mclust'
)

invisible(sapply(packages, library, character.only = TRUE))

# load data

dat <- readRDS("data_l1.rds")
dat_agg <- readRDS("data_l2.rds")

# Define variables: classclimate/wellbeing/behavioral problems/control for all analyzes
var_cc_l1 <- c("feess_ga", "feess_si", "soc_relatedness","soc_positivity", "sar_victim", "sar_perpetrator")
var_cc_l2 <- paste0(var_cc_l1, "_mean")
var_cc_l2_sd <- paste0(var_cc_l1, "_sd")

var_bp_l1 <- c("itrf_sw", "itrf_ad", "itrf_apd", "itrf_opp")
var_bp_l2 <- paste0(var_bp_l1, "_mean")
var_bp_l2_sd <- paste0(var_bp_l1, "_sd")

var_cont_l1 <- c("age", "grade", "sex", "migration_background")
var_cont_l2 <- paste0(var_cont_l1, "_mean")
var_cont_l2_sd <- paste0(var_cont_l1, "_sd")

# set parameters

knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# function for nicer output
html_table <- function(table, title = "", footnote = "") {
  title <- paste0("Table.<br><i>", title, "</i>")
  kable(table, caption = title) %>%
   kable_classic(full_width = FALSE) %>%
   footnote(footnote)
}

html_regression <- function(fit, ...) {
  sjPlot::tab_model(fit, ..., show.est = FALSE, show.std = TRUE, show.se = TRUE, show.ci = FALSE, string.est = "B", string.std = "Beta", string.se = "SE B", string.std_se = "SE", p.style = "stars") 
}

```

# Software acknowledgment

The following packages were used for this report:

`R` at version `r R.version$version.string` [@R-base].

```{r package_reference, results="asis", echo=FALSE}
wmisc::reference_package_version(packages)
```

------------------------------------------------------------------------

# Research questions

Q1: Is there a general factor that could be derived from various aspects of students relations within the classroom (ie. social inclusion; social position; bullying involvement; acceptance by the teacher) ? 

Q1a: On an individual level: Is it possible to conflate various aspects of students relations within the classroom into one distinct general factor (social-relational situation)?

Q1b: On a classroom level: Is it possible to conflate the classroom means of various aspects of students relations into on distinct general factor (social-relational classroom climate)?

Q2: What is the connection between students behavioral problems and various aspects of their social relations?

Q2a: On an individual level: To what extend can students behavioral problems in a class be explained by means of various aspects of their social relations?

Q2b: On a classroom level: To what extend can the average amount of behavioral problems in a classroom be explained by classroom means of various aspects of students social relations? 

Q2c: To what extend do classroom means of various aspects of students social relations contribute to individual behavioral classroom problems?


# Descriptives of all variables

```{r descriptives_l1}
.var <- c(var_cc_l1, var_bp_l1, "age", "sex_male", "migration_background_numeric")

dat %>% select(!!.var) %>% rename_items() %>% descriptives() %>% html_table

dat %>% group_by(grade) %>% summarise(n = n()) %>% html_table
dat %>% group_by(sex) %>% summarise(n = n()) %>% html_table
dat %>% group_by(migration_background) %>% summarise(n = n()) %>% html_table

dat %>% select(!!.var) %>% rename_items() %>% wmisc::nice_corrmatrix(upper = FALSE, numbered_columns = TRUE, char_p10 = "T") %>% html_table
```

```{r descriptives_l2}
.var <- c(var_cc_l2, var_bp_l2, "age_mean", "sex_male_mean", "migration_background_numeric_mean")
dat_agg %>% 
  select(!!.var) %>% 
  rename_items() %>% 
  descriptives() %>% 
  html_table()

dat_agg %>% 
  select(!!.var) %>% rename_items() %>% 
  wmisc::nice_corrmatrix(upper = FALSE, numbered_columns = TRUE, char_p10 = "T") %>% html_table
```


# Q1a

We conducted a parallel analysis to determine the number of factors to be extracted from the data. The analysis point to two factors (see figure xxx)

```{r parallel_l1, fig.cap="Parallel analysis scree plot."}
dat %>% select(!!var_cc_l1) %>% fa.parallel(main = "", fa = "fa")
```

```{r efa_l1}
dat %>% 
  select(!!var_cc_l1) %>% 
  rename_items() %>% 
  exploratory_fa(nfactors = 2, cut = 0.2, fm = "minres") %>% 
  html_table("Exploratory factor analyzes of wellbeing variables.")
```

# Q1b

We conducted a parallel analysis to determine the number of factors to be extracted from the data. The analysis point to two factors (see figure xxx).

```{r nfactors_l2, , fig.cap="Parallel analysis scree plot."}
dat_agg %>% 
  select(!!var_cc_l2) %>% 
  fa.parallel(main = "", fa = "fa")
```

```{r efa_l2}
dat_agg %>% 
  select(!!var_cc_l2) %>% 
  rename_items() %>% 
  exploratory_fa(nfactors = 2, fm = "minres") %>%
  html_table("Exploratory factor analyzes of classclimate variables.")
```

# Q2a

```{r}
dat_l12 <- full_join(dat, dat_agg, by = "id_class_teacher")
```


```{r}
#fit <- lme(itrf_sw ~ feess_si + feess_ga + soc_positivity + soc_relatedness + sar_victim + sar_perpetrator + age + migration_background + sex, random =~ 1 | id_class_teacher, data = dat, na.action = na.omit)

#html_regression(fit, title = "Social withdrawal regressed on wellbing variables")

fit_l1 <- lme(itrf_sw ~ feess_si + feess_ga + soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + migration_background + age + sex, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)

fit_l2 <- update(fit_l1, .~. +feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean)

html_regression(fit_l1, fit_l2, title = "Social withdrawal regressed on wellbing and classclimate variables", dv.labels = c("Model 1", "Model 2"))

```

```{r}
fit <- lme(itrf_ad ~ feess_si + feess_ga + soc_positivity  + soc_relatedness + sar_victim + sar_perpetrator+ age + migration_background + sex, random =~ 1 | id_class_teacher, data = dat, na.action = na.omit)
html_regression(fit, title = "Anxiety/depression regressed on wellbing variables")
```

```{r}
fit <- lme(itrf_apd ~ feess_si + feess_ga + soc_positivity  + soc_relatedness + sar_victim + sar_perpetrator+ age + migration_background + sex, random =~ 1 | id_class_teacher, data = dat, na.action = na.omit)
html_regression(fit, title = "Academic disorganized behavior regressed on wellbing variables")
```

```{r}
fit <- lme(itrf_opp ~ feess_si + feess_ga +  soc_positivity  + soc_relatedness + sar_victim + sar_perpetrator+ age + migration_background + sex, random =~ 1 | id_class_teacher, data = dat, na.action = na.omit)


html_regression(fit, title = "Oppositional behavior regressed on wellbing variables")
```

# Q2b

```{r}
fit <- lm(itrf_sw_mean ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean   + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean, data = dat_agg, na.action = na.omit)
html_regression(fit, title = "Social withdrawal regressed on classclimate")
```

```{r}
fit <- lm(itrf_ad_mean ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean + sar_victim_mean + sar_perpetrator_mean+ age_mean + migration_background_numeric_mean + sex_male_mean, data = dat_agg, na.action = na.omit)
html_regression(fit, title = "Anxiety/depression regressed on classclimate")
```

```{r}
fit <- lm(itrf_apd_mean ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean + sar_victim_mean + sar_perpetrator_mean+ age_mean + migration_background_numeric_mean + sex_male_mean, data = dat_agg, na.action = na.omit)
html_regression(fit, title = "Academic disorganized behavior regressed on classclimate")
```

```{r}
fit <- lm(itrf_opp_mean ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean, data = dat_agg, na.action = na.omit)
html_regression(fit, title = "Opositional behavior regressed on wellbing variables")
```

# Q2c

```{r}
dat_l12 <- full_join(dat, dat_agg, by = "id_class_teacher")
```


```{r}
fit <- lme(itrf_sw ~ feess_si + feess_ga + soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age + migration_background + age_mean + sex + migration_background_numeric_mean + sex_male_mean, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)
html_regression(fit, title = "Social withdrawal regressed on wellbing and classclimate variables")
```

```{r}
fit <- lme(itrf_ad ~ feess_si + feess_ga +  soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age + migration_background + age_mean + sex + migration_background_numeric_mean + sex_male_mean, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)
html_regression(fit, title = "Anxiety/Depression regressed on wellbing and classclimate variables")
```

```{r}
fit <- lme(itrf_apd ~ feess_si + feess_ga + soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age + migration_background + age_mean + sex + migration_background_numeric_mean + sex_male_mean, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)
html_regression(fit, title = "Academic disengagement regressed on wellbing and classclimate variables")
```

```{r}
fit <- lme(itrf_opp ~ feess_si + feess_ga + soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age + migration_background + age_mean + sex + migration_background_numeric_mean + sex_male_mean, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)
html_regression(fit, title = "Oppositional behavior regressed on wellbing and classclimate variables")
```




# Reference
