---
title: 'Analyses: Disentangling social-relational classroom climate and its relation
  to students'' behavioral problems'
author: "Jürgen Wilbert"
date: "Version (`r Sys.Date()`)"
output:
  html_document:
    theme: cosmo
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
bibliography: packages.bib
geometry: a4paper
csl: apa.csl
link-citations: yes
biblio-style: apalike
---

```{r setup, include=FALSE}
# load packages -----------------------------------------------------------

if (!("wmisc" %in% installed.packages())) 
  devtools::install_github("jazznbass/wmisc")
if (!("scaledic" %in% installed.packages())) 
  devtools::install_github("jazznbass/scaledic")

packages <- c(
  'sjPlot', 'knitr', 'semTable', 'kableExtra', 'bookdown', 
  'scaledic', 'Amelia', 'wmisc', 'psych', 'lavaan', 'nlme', 'lme4',
  'tidyverse', 'readxl', 'gridExtra', 'broom', 'sjPlot', 'mclust', 'huxtable'
)

invisible(sapply(packages, library, character.only = TRUE))

# load data

dat <- readRDS("data_l1.rds")
dat_agg <- readRDS("data_l2.rds")
dat_teachers <- readRDS("data_teachers.rds")
dat_alpha <- readRDS("data_alpha.rds")
# Define variables: classclimate/wellbeing/behavioral problems/control for all analyzes
var_cc_l1 <- c("feess_ga", "feess_si", "soc_relatedness","soc_positivity", "sar_victim", "sar_perpetrator")
var_cc_l2 <- paste0(var_cc_l1, "_mean")
var_cc_l2_sd <- paste0(var_cc_l1, "_sd")

var_bp_l1 <- c("itrf_sw", "itrf_ad", "itrf_apd", "itrf_opp")
var_bp_l2 <- paste0(var_bp_l1, "_mean")
var_bp_l2_sd <- paste0(var_bp_l1, "_sd")

var_cont_l1 <- c("age", "grade", "sex", "migration_background")
var_cont_l2 <- paste0(var_cont_l1, "_mean")
var_cont_l2_sd <- paste0(var_cont_l1, "_sd")

# set parameters

knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# function for nicer output
html_table <- function(table, title = "", footnote = "") {
  title <- paste0("Table.<br><i>", title, "</i>")
  kable(table, caption = title) %>%
   kable_classic(full_width = FALSE) %>%
   footnote(footnote)
}

html_regression <- function(fit, ...) {
  sjPlot::tab_model(fit, ..., show.est = FALSE, show.std = TRUE, show.se = TRUE, show.ci = FALSE, string.est = "B", string.std = "Beta", string.se = "SE B", string.std_se = "SE")#, p.style = "stars") 
}

poly_table <- function(x, caption = "", footnote = "", header = "", width = 1, ...) {
  
  x <- as_hux(x) %>%
    theme_article %>%
    set_caption(caption) 
  
  if (!identical(header, "")) {
      cols <- rep(" ", length(unlist(header)))
      for(i in seq_along(header)) cols[min(unlist(header[i]))] <- names(header)[i]
      x <- do.call(insert_row, c(list(ht = x, after = 0), as.list(cols)))
      for(i in seq_along(header)) {
        x <- do.call(merge_cells, list(x, row = 1, col = header[[i]]))
      }
      x <- x %>%
        set_align(1:2, everywhere, "center") %>% 
        set_tb_padding(1, everywhere, 0) %>% 
        set_bold(1, everywhere) %>%
        set_width(1)
  }
  
  if (!is.null(width)) x <- set_width(x, width)
  if (footnote != "") x <- add_footnote(x, footnote, italic = TRUE)
  
  x  
}
```

# Software acknowledgment

The following packages were used for this report:

`R` at version `r R.version$version.string` [@R-base].

```{r package_reference, results="asis", echo=FALSE}
wmisc::reference_package_version(packages)
```

------------------------------------------------------------------------

# Research questions

Q1: Is there a general factor that could be derived from various aspects of students relations within the classroom (i.e. social inclusion; social position; bullying involvement; acceptance by the teacher) ?

Q1a: On an individual level: Is it possible to conflate various aspects of students relations within the classroom into one distinct general factor (social-relational situation)?

Q1b: On a classroom level: Is it possible to conflate the classroom means of various aspects of students relations into on distinct general factor (social-relational classroom climate)?

Q2: What is the connection between students behavioral problems and various aspects of their social relations?

Q2a: On an individual level: To what extend can students behavioral problems in a class be explained by means of various aspects of their social relations?

Q2b: In extension of hypothesis Q2a: To what extend do classroom means of various aspects of students social relations contribute to individual behavioral classroom problems?

Q2c: On a classroom level: To what extend can the average amount of behavioral problems in a classroom be explained by classroom means of various aspects of students social relations?

# Analyzing strategy

We will start with a description of means, standard deviations, and correlations of all variables on an individual level and on a classroom aggregated level.

For addressing questions 1a (the factor structure of social-relation variables on an individual level) we will firstly determined the number of factors by conducting a parallel analysis. In a second step we will calculate an exploratory factor analysis (factor extraction method: minimal residuals). For addressing question 1b (the factor structure of social-relation variables on a classroom level) we will firstly calculate the means per class for all social-relation variables. Than we will proceed as for question 1b. The analyses for question 1a and b will be conducted with the psych package in R [@R-base; @R-psych].

For answering questions 2a, b, and c (the connection between behavioral problems and social relations), we will conduct a series of multilevel analyses (random intercept models with students nested in classrooms). For question 2a (individual level) we will set up models with each of the four behavioral problem variables as a criteria and the six social-relational variables as predictors. Additionally, we will add migration-background, age, and sex as additional predictors as these variables are expected to be correlated to behavioral problems. Sex and migration background (dichotomous variables) are contrast coded with -1 and 1 (Helmert contrast).

For question 2b (influence of social-relations on classroom level on individual behavior) we will firstly calculate the means of the six social-relations variables per class and add these variables as level 2 predictors. Additionally, we will also include the class means for the variables sex (proportion male students per class), migration background (proportion of migration background per class), and age (average age per class). Finally, all level 1 variables (as for question 2a) will be included as predictors.

For answering question 2c (influence of social-relations on classroom level on average behavorial problems in a classroom) we will set up a *mean-as-outcome* model for each of the for behavioral criteria. A *mean-as-outcome* model includes only level 2 predictors while the criteria is a level 1 variable (here: behavioral problems). As a result, the predictors will explain differences in the classroom means of the criteria variable. For this, we will include the mean social-relations variables and the mean age, the proportion migration background, and the the proportion male students as predictors (as in the models for questions 2b).

All estimators will be reported standardized. All analyses for questions 2a, b, and c will be conducted in R with the nlme package [@R-nlme]. Regression tables will be created with the help pf the sjplot package [@R-sjPlot].

The R code for all analyses and minimal data sets for reproducing the results will be provided in the supplementary of this paper.

# Material

```{r alpha}
  dat_alpha %>% html_table("Item Analyzes with the data of the present study.")
```

# Descriptives of all variables

```{r teachers_sample}
  dat_teachers %>% t() %>% html_table(title = "Teacher sample", footnote = "Age and active years as a teacher were administered as categorical variables (age: 20-30; 31-40, 41-50, 51-60, > 60; active years: < 1, 1-3, 4-10, > 10)")
```

```{r descriptives_l1}
.var <- c(var_cc_l1, var_bp_l1, "age", "sex_male", "migration_background_numeric")

dat %>% select(!!.var) %>% rename_items() %>% descriptives() %>% html_table

dat %>% group_by(grade) %>% summarise(n = n()) %>% html_table
dat %>% group_by(sex) %>% summarise(n = n()) %>% html_table
dat %>% group_by(migration_background) %>% summarise(n = n()) %>% html_table

dat %>% select(!!.var) %>% rename_items() %>% wmisc::nice_corrmatrix(upper = FALSE, numbered_columns = TRUE, char_p10 = "T") %>% html_table
```

```{r descriptives_l2}
.var <- c(var_cc_l2, var_bp_l2, "age_mean", "sex_male_mean", "migration_background_numeric_mean")
dat_agg %>% 
  select(!!.var) %>% 
  rename_items() %>% 
  descriptives() %>% 
  html_table()

dat_agg %>% 
  select(!!.var) %>% rename_items() %>% 
  wmisc::nice_corrmatrix(upper = FALSE, numbered_columns = TRUE, char_p10 = "T") %>% html_table
```

# Q1a

```{r parallel_l1, fig.cap="Parallel analysis scree plot."}
dat %>% select(!!var_cc_l1) %>% fa.parallel(main = "", fa = "fa")
```

```{r efa_l1}
dat %>% 
  select(!!var_cc_l1) %>% 
  rename_items() %>% 
  exploratory_fa(nfactors = 2, cut = 0.2, fm = "minres") %>% 
  html_table("Exploratory factor analyzes of social relation variables.")
```

# Q1b

We conducted a parallel analysis to determine the number of factors to be extracted from the data. The analysis point to two factors (see figure xxx).

```{r nfactors_l2, , fig.cap="Parallel analysis scree plot."}
dat_agg %>% 
  select(!!var_cc_l2) %>% 
  fa.parallel(main = "", fa = "fa")
```

```{r efa_l2}
dat_agg %>% 
  select(!!var_cc_l2) %>% 
  rename_items() %>% 
  exploratory_fa(nfactors = 2, fm = "minres") %>%
  html_table("Exploratory factor analyzes of mean social relation variables.")
```

# Q2a, Q2b, and Q2c

> Note for my co-authors: I changed the analyzes for Q2c from an aggregated OLS regression approach to a multilevel means-as-outcome model. While the unstandardized estimators are very similar in both models, the latter (means-as-outcome) model has much lower standardized betas. Probably this is due to a significant proportion of within class variance that has not been accounted for in the simple aggregated OLS regression (due to the aggregation). The change allows for representing all three models in one table, which I find much more convenient.

For calculating the influence of the average social relations within a class on the average behavior problems in a class, we fitted a mean-as-outcome model to the data. Mean-as-outcome model are multilevel models with level 2 predictors (here: the class level) but no level 1 predictors (here: the indivdual level). As a result, the predictors describe the changes in the grand mean of the criteria variable (here: behavior problems) that is due to compositional (here: means) aspects of students social relations within a class.

```{r}
dat_l12 <- full_join(dat, dat_agg, by = "id_class_teacher")
```

```{r}
fit_l1 <- lme(itrf_sw ~ feess_si + feess_ga + soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + migration_background + age + sex, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)

fit_l2 <- update(fit_l1, .~. +feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean)

fit_mean <- lme(itrf_sw ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean   + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean, random = ~  1|id_class_teacher, data = dat_l12, na.action = na.omit)

html_regression(fit_l1, fit_l2, fit_mean, title = "Social withdrawal regressed on individual social-relations and social-relational classclimate", dv.labels = c("Model 1 (L1)", "Model 2 (L1+L2)", "Model 3 (mean-as-outcome)"))
```

```{r}
fit_l1 <- lme(itrf_ad ~ feess_si + feess_ga + soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + migration_background + age + sex, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)

fit_l2 <- update(fit_l1, .~. +feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean)

fit_mean <- lme(itrf_ad ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean, random = ~  1|id_class_teacher, data = dat_l12, na.action = na.omit)

html_regression(fit_l1, fit_l2, fit_mean, title = "Anxiety/depression regressed on individual social-relations and social-relational classclimate", dv.labels = c("Model 1 (L1)", "Model 2 (L1+L2)", "Model 3 (mean-as-outcome)"))

```

```{r}
fit_l1 <- lme(itrf_apd ~ feess_si + feess_ga + soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + migration_background + age + sex, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)

fit_l2 <- update(fit_l1, .~. +feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean)

fit_mean <- lme(itrf_apd ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean   + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean, random = ~  1|id_class_teacher, data = dat_l12, na.action = na.omit)

html_regression(fit_l1, fit_l2, fit_mean, title = "Academic disorganized behavior regressed on individual social-relations and social-relational classclimate", dv.labels = c("Model 1 (L1)", "Model 2 (L1+L2)", "Model 3 (mean-as-outcome)"))

```

```{r}
fit_l1 <- lme(itrf_opp ~ feess_si + feess_ga + soc_relatedness + soc_positivity + sar_victim + sar_perpetrator + migration_background + age + sex, random =~ 1 | id_class_teacher, data = dat_l12, na.action = na.omit)

fit_l2 <- update(fit_l1, .~. +feess_si_mean + feess_ga_mean + soc_relatedness_mean + soc_positivity_mean + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean)

fit_mean <- lme(itrf_opp ~ feess_si_mean + feess_ga_mean + soc_positivity_mean + soc_relatedness_mean   + sar_victim_mean + sar_perpetrator_mean + age_mean + migration_background_numeric_mean + sex_male_mean, random = ~  1|id_class_teacher, data = dat_l12, na.action = na.omit)

html_regression(fit_l1, fit_l2, fit_mean, title = "Oppositional behavior regressed on individual social-relations and social-relational classclimate", dv.labels = c("Model 1 (L1)", "Model 2 (L1+L2)", "Model 3 (mean-as-outcome)"))

```

# Reference
